---
title: 'Bibliometric Analysis: Matthew A. Peeples'
author: "Matthew A. Peeples"
date: "2025-02-15"
output: pdf_document
header-includes:
    \usepackage{helvet}
  \renewcommand{\familydefault}{\sfdefault}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document contains a set of bibliometric and citation analyses for the published work of Matthew Peeples. This information was prepared for the purposes of consideration for academic promotion. The data used here come directly from Google Scholar and Scopus and were scraped (2/15/2025) using the Harzing Publish or Perish Windows application and the R package 'scholar'. All of the data and raw code used in this document can be viewed on [Github here: https://github.com/mpeeples2008/PeeplesCitation](https://github.com/mpeeples2008/PeeplesCitation). Note that I have been at ASU since 2015 and received tenure and was promoted in 2019 (submitted review in 2018). These data are current up to 2024 which is the last full year available at the time of writing (2/15/2025).

## Google Scholar Citations Per Year

The plot below shows the number of citations I have received per year up to 2024, with the pre- and post-tenure intervals indicated. Note that the number of annual citations has generally shown an upward trajectory. The large peak in 2023 is associated with my co-edited *Oxford Handbook of Archaeological Network Science* which included 42 chapters explicitly focused on archaeological networks generating a large number of citations. Although the book was officially released in 2024 the online version was first made available in December 2023 and thus, Google Scholar counts those citations in 2023. 

As of 2/15/2025 my work has been **cited 2,945 times** with an **h-index of 30** and an **i10-index of 51**. At the time of my last review for tenure I had received **717 citations** with a **h-index of 15**. This suggests that citations and apparent impact have both increased substantially since tenure.

```{r cites, message=F, warning=F, echo=F, fig.height=3}
library(scholar)
## Define the id for Matthew Peeples
id <- '8B39PWQAAAAJ'

## Get citation history
ct <- get_citation_history(id)
ct <- ct[-nrow(ct),] # remove current year


## Plot citation trend
library(ggplot2)
p1 <- ggplot(ct, aes(year, cites)) + geom_line() + geom_point(size=4) +
  scale_x_continuous("Year", labels = as.character(ct$year), breaks = ct$year) +
  ylab("Citations in Google Scholar") +
  theme_classic() +
  theme(axis.text=element_text(size=10), axis.title=element_text(size=10))

p1 <- p1 + annotate('rect', xmin=2018.5, xmax=2024, ymin=0, ymax=500,
                    alpha=0.2, fill='blue') + annotate('text', x=2021,y=50, label='Post-Tenure', size=4, col='blue')

p1 <- p1 + annotate('rect', xmin=2015, xmax=2018.5, ymin=0, ymax=500,
                    alpha=0.2, fill='red') + annotate('text', x=2017,y=50, label='Pre-Tenure', size=4, col='red')

p1

```

\newpage
## Journal Impact Factors

Next, I explore the Scopus impact factors/percentiles for journals in which I have published through time. In order to do this, I created a file of all of my journal publications that have Scopus CiteScores and recorded both the CiteScore and the highest percentile by field/area for each journals. The vast majority of my publications through time have appeared in highly ranked journals. All but one of my publications are in journals in the **top quartile** by field and **84% are in the top 10% of journals by field**. Given the consistency with which I have published in highly ranked journals there is little room for improvement through time but, as the boxplots below illustrate, there is a trend toward my publications narrowing to the most highly ranked journals by field in my time at ASU and post-tenure.

\
\


```{r impact, echo=F, warning=F, message=F}

library(ggplot2)
impact <- read.csv('impact.csv',header=T)

impact$Period <- factor(impact$Period, levels = c('pre-ASU','pre-tenure','post-tenure'))

ggplot(data=impact, aes(y=Percentile, x=Period)) + 
  geom_boxplot()+
  theme_bw()+
  theme(axis.text=element_text(size=10), axis.title=element_text(size=10))


```




\newpage
## Citations and Publication Reach

In order to document the expanding reach of my research since I earned tenure, I scraped and downloaded information on all papers that have cited my work using Publish or Perish since my first publication in 2003. I then cleaned and checked these data (in particular adding years where they were absent) and added two additional variables: Field and Language. These represent the general academic field in which a citing publication falls and the original language in which it was published. Languages were automatically detected using the Google "Language Detect" tool. Academic field was determined based on the primary field for a given publication in Scopus and where this was not available, based on my own assessment of the publication content or title.

The plots below show the proportion of publications per year (since I arrived at ASU) that fall into academic fields outside of anthropology and archaeology (top) and publications appearing in languages other than English (bottom). Note that there are pronounced increases in both non-Anthropology/Archaeology citations and non-English citations in recent years, demonstrating the increasing reach of my research.
\

```{r reach, message=F, warning=F, echo=F, fig.height=6}
library(ggplot2)
library(ggpubr)

cites <- read.csv('PoPCites_updated.csv', header=T)
cites <- unique(cites[,1:6])

dat <- cites[,which(colnames(cites) %in% c('Field','Year','Language'))]

out <- as.data.frame(matrix(0,10,3))
colnames(out) <- c('Year','OutsideCitesPerYear','Language')
out[,1] <- seq(2015,2024)
for (i in out[,1])  {
  temp <- dat[which(dat$Year==i),]
  out[i-2014,2] <- length(which(!temp$Field  %in% c('Anthropology','Archaeology')))/nrow(temp)
  out[i-2014,3] <- length(which(!temp$Language  %in% c('English')))/nrow(temp)
}


p1 <- ggplot(data=out) +
  geom_line(aes(x=Year,y=OutsideCitesPerYear)) +
  scale_x_continuous("Year", labels = as.character(out$Year), breaks = out$Year) +
  ylab("Propotion of Citations Outside of Anthropology and Archaeology") +
  theme_classic() +
  theme(axis.text=element_text(size=7), axis.title=element_text(size=7))


p1 <- p1 + annotate('rect', xmin=2019, xmax=2024, ymin=0.14, ymax=0.275,
    alpha=0.2, fill='blue') + annotate('text', x=2021,y=0.157, label='Post-Tenure', size=7, col='blue')



p2 <- ggplot(data=out) +
  geom_line(aes(x=Year,y=Language)) +
  scale_x_continuous("Year", labels = as.character(out$Year), breaks = out$Year) +
  ylab("Propotion of Citations in Non-English Language Publications") +
  theme_classic() +
  theme(axis.text=element_text(size=7), axis.title=element_text(size=7))


p2 <- p2 + annotate('rect', xmin=2019, xmax=2024, ymin=0, ymax=0.075,
             alpha=0.2, fill='blue') + annotate('text', x=2021,y=0.01, label='Post-Tenure', size=8, col='blue')

ggarrange(p1,p2, nrow=2, ncol=1)
```

\newpage
In order to dig into this a little further, I display a chart with the fields (other than Anthropology and Archaeology) for publications citing my work. As this shows, my work is being recognized and used by researchers in the computational, social, and environmental sciences in particular. I have received a few additional citations in fields like education, health science, library science, cognitive science, mathematics, and political science largely based on methodological contributions of my research. 
\

```{r field, warning=F, message=F, echo=FALSE,fig.height=3}
field <- as.data.frame(table(dat$Field))
field <- field[-which(field$Var1 %in% c('Anthropology','Archaeology','archaeology','')),]
field2 <- field[which(field$Freq>14),]
field2[11,2] <-sum(field$Freq[which(field$Freq<15)])
levels(field2$Var1) <- c(levels(field2$Var1), 'Other')
field2[11,1] <- 'Other'
colnames(field2) <- c('Field','Freq')

field2$fraction = field2$Freq / sum(field2$Freq)
field2$ymax = cumsum(field2$fraction)
field2$ymin = c(0, head(field2$ymax, n=-1))

# Make the plot
p2 <- ggplot(field2, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Field)) +
  geom_rect() +
  scale_fill_brewer(palette='Spectral') +
  coord_polar(theta="y") + 
  xlim(c(2, 4)) +
  theme_minimal()

p2 <- p2 +  theme(axis.ticks=element_blank(),  
           axis.text.y=element_blank(), 
           axis.title=element_blank())  
p2
```


The chart below shows the languages of non-English publications that cite my work. As of 2024, my work has been cited in 18 different languages, in particular in Spanish and French. The "other" languages representing only one citation each are Arabic, Czech, Dutch, German, Indonesian, Norwegian, Polish, Serbian, and Swedish. This illustrates that my publications are beginning to be recognized broadly across the world.
\

```{r languages, echo=F,fig.height=3}
lang <- as.data.frame(table(dat$Language))
lang <- lang[-which(lang$Var1=='English'),]
lang2 <- lang[which(lang$Freq>1),]
lang2[9,2] <-sum(lang$Freq[which(lang$Freq==1)])
levels(lang2$Var1) <- c(levels(lang2$Var1), 'Other')
lang2[9,1] <- 'Other'
colnames(lang2) <- c('Language','Freq')

lang2$fraction = lang2$Freq / sum(lang2$Freq)
lang2$ymax = cumsum(lang2$fraction)
lang2$ymin = c(0, head(lang2$ymax, n=-1))

# Make the plot
p1 <- ggplot(lang2, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Language)) +
  geom_rect() +
  scale_fill_brewer(palette='Spectral') +
  coord_polar(theta="y") + 
  xlim(c(2, 4)) +
  theme_minimal()

p1 <- p1 +  theme(axis.ticks=element_blank(), 
           axis.title=element_blank(),  
           axis.text.y=element_blank()) 

p1
```

\newpage
## Co-Author Networks

One additional way of looking at the impact and reach of my publications is exploring my co-authorship network. In order to do this, I scraped all of my "coauthor" data from Google Scholar using the 'scholar' package. Note that this package will only locate co-authors who have an active Google Scholar profile. The resulting data are shown below as an ego network focused on me. Note that there are many distinct sub-groups in this network that are only connected via their path through me. Indeed, the transitivity of this network is relatively low at at **transitivity = 0.105** indicating a ego centered network with substantial structural holes (gaps in network structure). Although this sample is somewhat limited, the results here illustrate my key role as a "broker" and collaborator connecting many communities of scholars who would otherwise be disconnected.

```{r network, echo=F, warning=F, message=F, fig.height=7}

library(scholar)

## Define the id for Matthew Peeples
id <- '8B39PWQAAAAJ'

coauthor_network <- get_coauthors('8B39PWQAAAAJ', n_coauthors = 500)

coauthor_network <- coauthor_network[-which(coauthor_network[,2] %in% c('Search Help','About Scholar')),]

plot_coauthors(coauthor_network, size_labels=1)

```

\newpage
## International Collaborations 

In order to illustrate my international profile as a researchers I provide a map showing the worldwide distribution of my collaborators. As this map shows, I have collaborators across the US as well as many in Europe. This map was generated by manually compiling the cities/affiliation for my co-authors, including only the interval since I started my position at ASU.


```{r analytics, warning=F, message=F, echo=F, fig.height=3}
library(maps)
library(tidyverse)
library(ggthemes)

world_coordinates <- map_data("world") 
collab <- read.csv("collabs.csv",header=T) 
  
ggplot() + 
  geom_map( 
    data = world_coordinates, map = world_coordinates, 
    aes(long, lat, map_id = region), 
    color = "lightblue", fill= "lightyellow"
  )+ 
geom_point( 
    data = collab, 
    aes(x=long, y=lat), 
    alpha = 0.5, size=2 
  ) + 
  theme_minimal() +
  theme(legend.position='none')

```


In addition to looking at my publications, it is also instructive to look to the use of online resoruces I have created. For example, accordding to Google Analytics data scraped using the 'googleAnalyticsR' package the online companion to my book *Network Science in Archaeology* available at [https://book.archnetworks.net](https://book.archnetworks.net) has received **6,823 unique visitors** since the book was released which come from more than 100 different countries or regions. The table below shows information on unique visits and active users by country for the top 20 counties represented. This list shows that my work is being used by researchers across many regions of the world.

```{r analytics2, warning=F, message=F, echo=F}
library(googleAnalyticsR)
library(maps)

ga_auth(email="mpeeple@asu.edu")

my_property_id <- 319575878
from_date <- "2023-03-01"
to_date <- "2025-02-15"
           
overall <- ga_data(
  my_property_id,
  metrics = c("activeUsers", "newUsers","sessions", "screenPageViews"),
  date_range = c(from_date, to_date)
  )

country <- ga_data(
  my_property_id,
  dimensions = c("country"),
  metrics = c("activeUsers", "newUsers","sessions", "screenPageViews"),
  date_range = c(from_date, to_date)
)

country <- as.data.frame(country)

country[1:20,]
```

